Reconstructing the last known movements of one of Nature's giants

### 2017-06-23 manuscript/blue-whale-ms.tex
Add CT edits to abstract

\paragraph{Understanding animal movements is crucial for effective conservation, especially as species ranges shift due to climate change\cite{runge2014conserving,robinson2009travelling}. 
Current methods for monitoring movement behaviours are often expensive, and do not provide information on historical or long-term range movements.
Animal tissues housed in museum collections present an alternative, providing biochemical records of behaviour that can be used to reconstruct historical individual-level movements relatively inexpensively\cite{newsome2010using}. 
Here we combine stable isotope analyses of baleen with novel agent-based simulation models, to reconstruct the last years of an iconic blue whale (Balaenoptera musculus) that stranded off Ireland in 1891 and is now on display at the Natural History Museum, London (NHM). 
We recover individual-level movement information at high temporal resolution over six years and show distinct periods of foraging and short distance north-south migrations in sub-Arctic waters and calving in subtropical waters. 
We also detect isotopic signatures consistent with pregnancy and lactation in the year before the whale's death. 
These results supply the first quantitative evidence of historical migrations in blue whales, and suggest that blue whale residency and movement patterns in the North Atlantic have remained similar for hundreds of years, despite changing human pressures in the region. 
Historic tissue collections can provide unique insight into individual-level movement behaviours associated with changing climatic, ecological conditions and human-induced pressures, and reveal aspects of life history. 
Combining biochemical tracers with simulation modeling unlocks the information recorded in these irreplaceable archives.}

>>>>

\paragraph{Understanding animal movements is crucial for effective conservation, especially as species ranges shift due to climate change\cite{runge2014conserving,robinson2009travelling}. 
Animal tissues housed in museum collections provide biochemical records of behaviour that can be used to reconstruct individual-level movements over long time scales and under historic climatic and ecological conditions\cite{newsome2010using}. 
Here we combine stable isotope analyses of baleen with novel agent-based simulation models, to reconstruct movements over the last six years of the life of an iconic blue whale (\textit{Balaenoptera musculus}) that stranded off Ireland in 1891 and is now on display at the Natural History Museum, London (NHM). 
We identify two distinct phases in the whale's behaviour: repeated annual north-south migrations in sub-Arctic waters, and residency in subtropical waters. 
Subtropical residency is associated with isotopic signatures consistent with pregnancy and lactation in the year before the whale’s death. 
These results supply the first quantitative multi-yeat records of historical migrations in individual blue whales, and suggest that blue whale residency and movement patterns in the North Atlantic have remained similar for hundreds of years, despite changing human pressures in the region. 
Historic tissue collections can provide unique insight into individual-level movement behaviours associated with changing climatic, ecological conditions and human-induced pressures, and reveal aspects of life history. 
Combining biochemical tracers with simulation modelling unlocks the information recorded in these irreplaceable archives.}
